Appendix

### 2025-04-09 manuscript_rmarkdown/manuscript_appendix.Rmd
Rearrange appendix sections.

# Appendix C: Simulations with other optimisation criteria

Invariance of location- and scale-invariant power transformations was assessed using the optimisation criteria in [Appendix B](Appendix B: Optimisation of transformation parameters).
This follows the simulation in the main manuscript, where MLE was used for optimization.
In short, we first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
The second distribution was a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
The third distribution was a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution using different optimisation criteria.
To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
Similarly, to assess scale invariance, each distribution was multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.

@@

# Appendix D: Simulations on clean data

@@

# Appendix D: Experimental results

@@

Results for Box-Cox transformations are shown in Tables

@@

\begin{table}
\begin{center}
\caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality. 
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
\label{tab:experimental-results-appendix-residuals}
\begin{tabular}{l | r r r r r-bc}

@@

# Appendix F: Normalisation before transformation

Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors after transformation to normality of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
Table \ref{tab:normalisation-before-transformation-lambda-appendix} shows the corresponding values for transformation parameter $\lambda$, 
\ref{tab:normalisation-before-transformation-p-value-appendix} shows p-values of the empirical central normality test.
Normalisation prior to transformation prevents unexpectedly high $\lambda$ values in the presence of large location shifts, as can be observed for the latitude feature.
Normalisation prior to transformation, however, does not consistently improve residual errors after transformation. 
Interestingly, use of normalisation prior to transformation with robust power transformation using the method by Raymaekers and Rousseeuw [@Raymaekers2024-zf] led to higher residual errors for arterial wall thickness, fuel efficiency and latitude features.
This is similar to observations by @Raymaekers2024-zf, who observed that their method can emphasize values of outliers after transformation.
Location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.
Not surprisingly, normalisation before transformation does not noticeably improve central normality, as assessed using the empirical central normality test.

data <- .assess_standardisation_before_transformation()
data <- data[, mget(c("residuals_central", "lambda", "p_value", "transformation", "dataset"))]
data[, "residuals_central" := round(residuals_central, digits = 1L)]
data[, "lambda" := round(lambda, digits = 1L)]
data[, "p_value" := round(p_value, digits = 3L)]
data.table::dcast(data, transformation ~ dataset, value.var = "residuals_central")
data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
```

>>>>

## Simulations with other optimisation criteria

Invariance of location- and scale-invariant power transformations was assessed 
using the optimisation criteria in [Appendix B](Appendix B: Optimisation of transformation parameters).
This follows the simulation in the main manuscript, where MLE was used for optimization.
In short, we first randomly drew $10000$ values from a normal distribution:
$\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$,
or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
The second distribution was a right-skewed normal distribution 
$\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
The third distribution was a left-skewed normal distribution 
$\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.

We then computed transformation parameter $\lambda$ using the original definitions
(equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) 
and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} 
and \ref{eqn:yeo-johnson-invariant}) for each distribution using different optimisation criteria.
To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
Similarly, to assess scale invariance, each distribution was multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.

@@

# Appendix E: Simulations on clean data

@@

# Appendix F: Experimental results

@@

Results for Box-Cox transformations are shown in Tables \ref{tab:experimental-results-appendix-residuals-bc},
\ref{tab:experimental-results-appendix-lambda-bc} and \ref{tab:experimental-results-appendix-p-value-bc}.

@@

\begin{table}
\begin{center}
\caption{Residual errors for features from real-world datasets after Box-Cox transformation to normality. 
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling. 
AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
\label{tab:experimental-results-appendix-residuals-bc}
\begin{tabular}{l | r r r r r}

@@
